\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2018-01-30 eg6-letters/NIM/EG6_RTPC_NIM.tex
Minor fixes on NIM for resubmission

The original design of the RTPC was developed for the BoNuS 
experiment at Jefferson Lab which took data with CLAS in 
2005~\cite{Fenker:2008zz}. Significant improvements were made to the RTPC mechanical 
structure and fabrication technique that both increased the acceptance and 
reduced the amount of material in the path of the outgoing particles. 
Moreover, the data aquisition electronic was improved to increase the event 
readout rate. The enhanced design, 
used in the 2009 DVCS experiment, is described in section \ref{sec_design} of 
this article. The data acquisition system is described in section \ref{sec_readout}, 
the calibration methods in section \ref{sec_calib} and the tracking algorithm in section
\ref{sec_tracking}. Finally, the overall performances
of the RTPC are described in section \ref{sec_perfor}.

@@

We detail here the different regions shown in Figure~\ref{fig:RTPC_1_4} 
starting from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6~atm $^4$He target extends along the beamline forming the detector 
     central axis. It is a 6~mm diameter Kapton straw with a 27~\textmu{}m wall 
     of 292~mm length such that its entrance and exit 15~\textmu{}m aluminum windows 
     are placed outside of the detector volume. The detector and the target are 
     placed in the center of the solenoid, 64~cm upstream of the CLAS center.
   \item The first gas gap covers the radial range from 3~mm to 20~mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 \textmu{}m thick window made of grounded aluminized Mylar.
   \item The second gas gap region extends between 20~mm and 30~mm and is 
      filled with the gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether 
      (DME).  This region is surrounded by a 4~\textmu{}m thick window made of 
      aluminized Mylar set at $-4260$~V to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and 
      extends from the cathode to the first GEM, 60~mm away from the beam axis.  
      The electric field in this region is perpendicular to the beam and 
      averages around 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66~mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600~V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charges 
      after they have been multiplied by the GEMs. Pre-amplifiers are plugged 
      directly on its outer side and transmit the signal to the data 
      acquisition electronics.
\end{itemize}

@@

In order to experimentally determine the value of $T_{max}$, we analyse the
time profile of hits from identified good tracks, as presented in 
Figure~\ref{fig:TDC_profile}. We can clearly observe the expected dropping 
edge and define $T_{max}$, as the time at which the distribution passes 
below half the maximum number of hits in the histogram. This value was 
measured in bins along the 200~mm RTPC's length to take into account 
variations in the electric and magnetic field. We show this result for one 
experimental run in Figure~\ref{fig:RunNumber_61551_TDCmax_Zslice}.

@@

\subsection{Energy Loss Corrections}\label{sec_eloss}
Energy loss between the target and drift region was significant in the RTPC and 
necessitated a correction for optimal momentum reconstruction at the primary 
interaction vertex.  The dominant loss was in the 27~\textmu{}m thick Kapton target 
wall, with significant contributions also from the pressurized target gas and 
the foils before reaching the drift region.  Corrections were developed based 
on GEANT4 simulations with the full RTPC geometry and parameterized in terms of 
recoil curvature in the drift region and polar angle, separately for all recoil 
hypotheses ($p$, $d$, $^3$H, $^3$He, $^4$He).  At our average coherent $^4$He 
DVCS kinematics, energy losses were about 5 MeV, while for $e-^4$He elastic 
scattering losses were about 3 MeV, which corresponds to momentum corrections 
of 25\% and 15\%, respectively.

>>>>

The original design of the RTPC was developed for the BoNuS 
experiment at Jefferson Lab which took data with CLAS in 
2005~\cite{Fenker:2008zz}. Significant improvements were made to the RTPC mechanical 
structure and fabrication technique that both increased the acceptance and 
reduced the amount of material in the path of the outgoing particles. 
Moreover, the data acquisition electronic was improved to increase the event 
readout rate. The enhanced design, 
used in the 2009 DVCS experiment, is described in section \ref{sec_design} of 
this article. The data acquisition system is described in section \ref{sec_readout}, 
the calibration methods in section \ref{sec_calib} and the tracking algorithm in section
\ref{sec_tracking}. Finally, the overall performances
of the RTPC are described in section \ref{sec_perfor}.

@@

We detail here the different regions shown in Figure~\ref{fig:RTPC_1_4} 
starting from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6~atm $^4$He target extends along the beamline forming the detector 
     central axis. It is a 6~mm diameter Kapton straw with a 27~\textmu{}m wall 
     of 292~mm length such that its entrance and exit 15~\textmu{}m aluminum windows 
     are placed outside of the detector volume. The detector and the target are 
     placed in the center of the solenoid, 64~cm upstream of the CLAS center.
   \item The first gas gap covers the radial range from 3~mm to 20~mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 \textmu{}m thick window made of grounded aluminized Mylar.
   \item The second gas gap region extends between 20~mm and 30~mm and is 
      filled with the gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether 
      (DME).  This region is surrounded by a 4~\textmu{}m thick window made of 
      aluminized Mylar set at $-4260$~V to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and 
      extends from the cathode to the first GEM, 60~mm away from the beam axis.  
      The electric field in this region is perpendicular to the beam and 
      averages around 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66~mm. The first GEM layer is 
      set to $\Delta V=1620$~V relative to the cathode foil and then each subsequent 
      layer is set to a lower voltage relative to the previous to obtain a 
      strong (~1600~V/cm) electric field between the GEM foils. A 275~V bias is 
      applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charges 
      after they have been multiplied by the GEMs. Pre-amplifiers are plugged 
      directly on its outer side and transmit the signal to the data 
      acquisition electronics.
\end{itemize}

@@

In order to experimentally determine the value of $T_{max}$, we analyze the
time profile of hits from identified good tracks, as presented in 
Figure~\ref{fig:TDC_profile}. We can clearly observe the expected dropping 
edge and define $T_{max}$, as the time at which the distribution passes 
below half the maximum number of hits in the histogram. This value was 
measured in bins along the 200~mm RTPC's length to take into account 
variations in the electric and magnetic field. We show this result for one 
experimental run in Figure~\ref{fig:RunNumber_61551_TDCmax_Zslice}.

@@

\subsection{Energy Loss Corrections}\label{sec_eloss}
Energy loss between the target and drift region was significant in the RTPC and 
necessitated a correction for optimal momentum reconstruction at the primary 
interaction vertex.  The dominant loss was in the 27~\textmu{}m thick Kapton target 
wall, with significant contributions also from the pressurized target gas and 
the foils before reaching the drift region.  Corrections were developed based 
on GEANT4 simulations with the full RTPC geometry and parametrized in terms of 
recoil curvature in the drift region and polar angle, separately for all recoil 
hypotheses ($p$, $d$, $^3$H, $^3$He, $^4$He).  At our average coherent $^4$He 
DVCS kinematics, energy losses were about 5 MeV, while for $e-^4$He elastic 
scattering losses were about 3 MeV, which corresponds to momentum corrections 
of 25\% and 15\%, respectively.
